Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-02-07 Ben_Manuscripts/structure_paper/Draft.tex
Updates to intro

%MRS5: want to get people engaged from the beginning. The first
%paragraph should describe a problem to be solved. Then you can get
%into the motivating details (like background of the alternatives, and
%why we need something different).  I feel like this is an
%introduction into different types of membranes, though it should be
%an introduction to the problem that this paper takes steps towards
%solving.

  Problem to be solved

  Reverse osmosis (RO) and nanofiltration (NF) are membrane-based separation
  processes widely used to create potable or reusable water by removing salts and
  other dissolved compounds from water sources. RO membranes are typically thin
  film composite (TFC) membranes with a porous support layer and a dense
  crosslinked polymer active layer which near indiscriminately 
%MRS5: 
%  Maybe ``impermeable to all solutes?'' or ``separates all solutes from water.''  As written, it's not stated what the solutes are separated from.
  separates all solutes. 
NF membranes have porous architectures with pores on the order of 1 nm
  in size which separate primarily based on solute size
  \cite{van_der_bruggen_review_2003}.

  We need membranes that are more selective than RO and NF. Recovery of valued
  products from complex aqueous and organic waste is of increasing interest. For
  example, flowback water (FW) produced during hydraulic fracturing is a complex
  wastewater full of potentially valuable dissolved organic compounds
  \cite{dischinger_application_2017}. There is increasing pressure to reuse
  hydraulic fracturing water rather than dispose of it in order to reduce social
  and envrionmental impacts as well as cost \cite{theodori_hydraulic_2014}.
  Rather than dispose of the wastestream generated in the recycling process, we
  can instead use highly selective membranes in order to successfully recover the
  useful compounds.

@@

@@

  The feasiblity of LLC membranes has been demonstrated using LLCs that form
  the type 1 bicontinuous cubic (Q\textsubscript{I})
  \cite{hatakeyama_water_2011,hatakeyama_nanoporous_2010,carter_glycerol-based_2012},
  and the inverted hexagonal (H\textsubscript{II}) \cite{zhou_supported_2005}
  phases (See Figure~\ref{fig:bcc_v_hII}). When separating organic solutes from
  NaCl, Q\textsubscript{I} phase membrane filtration experiments have shown
  selectivity 2--3 times higher than commercial RO and 6--12 times higher than
  commercial NF membranes \cite{dischinger_application_2017}.  When separating a
  series of various sized dyes, the H\textsubscript{II} phase membrane showed
  complete rejection of dyes bigger than 1.2 nm in size
  \cite{zhou_supported_2005}.

%MRS5: be more explicit about current efforts. Also use the
%opportunity of a thesis statement to convince the readership of
%something.  what do you want to convince them?
  We will focus our current efforts around the H\textsubscript{II} phase since
  its geometry should give optimal transport performance compared to the
  Q\textsubscript{I} phase \cite{matyka_tortuosity-porosity_2008}. The
  H\textsubscript{II} phase forms at room temperature in the presence of c.a.~10
  wt\% water and consists of hexagonally packed, straight, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in literature, has been
  referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014} (See Figure~\ref{fig:hII_phase}).  These
  hexagonal columnar phases represent the ideal geometry for membrane transport.

@@

%MRS5: not a strong thesis.  Why has it been revived?  The reasons for that would be a better thesis.
  Research into the H\textsubscript{II} phase of polymerized LLC membranes has
  been revived in recent years. During early stages of exploration, hexagonal
  mesophases formed by Na-GA3C11 (Figure~\ref{fig:assembly}) could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. For this reason, research efforts
  focused on the Q\textsubscript{I} phase, whose geometry does not require
  alignment. In 2014, Feng et al.~showed that one can align 
  Col\textsubscript{h} mesophases using a magnetic field with subsequent
  cross-linking to lock the structure in place \cite{feng_scalable_2014}. In
  2016, Feng et al. showed that one could obtain the same result using a second
  technique termed soft confinement \cite{feng_thin_2016}. Current efforts are
  focused on extending the method to the H\textsubscript{II} phase and 
  characterizing the performance of these newly aligned systems.

@@

%MRS5: this is a better thesis. This sentence sets out a problem to the solved.
  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to Na-GA3C11 with some characterization done after minor structural
  modifications. Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing
  \cite{resel_structural_2000}. In a later study of rejection performance, it was
  shown that membranes formed by cross-linked Na-GA3C11 in the
  H\textsubscript{II} phase cannot separate solutes less than 1.2 nm in diameter
  because the pores are too large \cite{zhou_supported_2005}. We do not yet
  understand how to controllably reduce the effective pore size or how to tune
  the chemical environment in the nanopores of this or related materials for
  small molecule separations. The only source of predictive modeling for
  LLC systems have been macroscopic models that likely do not adequately describe
  transport at these length scales \cite{hatakeyama_water_2011}. It will be
  challenging to efficiently narrow down the large design space in a laboratory
  setting without a robust model.

@@

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport in these ordered, nanoporous
  organic systems, we must first gain a thorough understanding of the nanoscopic
  structure of LLC polymer membranes. 
%MRS5: repeated use of model below, both constructing and developing?
Our approach to constructing a general
  model will follow the development of a model of the assembly formed by
  Na-GA3C11 since it has sufficient experimental characterization. 
%MRS5: maybe clean up/motivate abit more about why Col_h is worth starting with.
We have also
  narrowed our scope to the development of a model of the Col\textsubscript{h}
  phase membrane. Compared to the H\textsubscript{II} phase, the
  Col\textsubscript{h} phase is a simpler starting point, due to the absence of
  water, and has detailed experimental wide-angle X-ray scattering (WAXS)
  patterns useful for reconstructing structural data. The two phases appear to
  share similar structural characteristics since the pore spacings in each system
  are in close agreement \cite{feng_thin_2016,resel_structural_2000}.

  Despite having structural data, there is still information which experiment
  cannot definitively answer. 
%MRS5: does a level of understanding suggest, or just is the level of understanding those facts?  and are there citations for these? are they inferred by the existing x-ray data.
Our current level of understanding suggests that
  monomers partition into layers with head groups facing towards the center, then 
  stack on top of each other to create pores that pack together hexagonally.
  There are several key questions that we will investigate in order to
  expand this understanding:

>>>>

  %MRS5: want to get people engaged from the beginning. The first paragraph
  %should describe a problem to be solved. Then you can get into the motivating
  %details (like background of the alternatives, and why we need something
  %different).  I feel like this is an introduction into different types of
  %membranes, though it should be an introduction to the problem that this paper
  %takes steps towards solving.

  A highly selective membrane can be useful for the recovery of valuable products
  from complex aqueous and organic waste.
  %We need membranes that are more selective than RO and NF. Recovery of valued
  %products from complex aqueous and organic waste is of increasing interest. 
  For example, flowback water (FW) produced during hydraulic fracturing is a complex
  wastewater full of potentially valuable dissolved organic compounds
  \cite{dischinger_application_2017}. There is increasing pressure to reuse
  hydraulic fracturing water rather than dispose of it in order to reduce social
  and environmental impacts as well as cost \cite{theodori_hydraulic_2014}.
  Rather than dispose of the wastestream generated in the recycling process, we
  can instead use highly selective membranes in order to successfully recover
  useful compounds.

@@

% BJC: Maybe I don't even need to dive into explaining RO and NF.
%  Reverse osmosis (RO) and nanofiltration (NF) are membrane-based separation
%  processes widely used to create potable or reusable water by removing salts and
%  other dissolved compounds from water sources. RO membranes are typically thin
%  film composite (TFC) membranes with a porous support layer and a dense
%  crosslinked polymer active layer which  separates all solutes from water. NF
%  membranes have porous architectures with pores on the order of 1 nm in size
%  which separate primarily based on solute size
%  \cite{van_der_bruggen_review_2003}.

@@

  The feasibility of LLC membranes has been demonstrated using LLCs that form
  the type 1 bicontinuous cubic (Q\textsubscript{I})
  \cite{hatakeyama_water_2011,hatakeyama_nanoporous_2010,carter_glycerol-based_2012},
  and the inverted hexagonal (H\textsubscript{II}) \cite{zhou_supported_2005}
  phases (See Figure~\ref{fig:bcc_v_hII}). When separating organic solutes from
  NaCl, Q\textsubscript{I} phase membrane filtration experiments have shown
  selectivity 2--3 times higher than commercial RO and 6--12 times higher than
  commercial NF membranes \cite{dischinger_application_2017}.  When separating a
  series of various sized dyes, the H\textsubscript{II} phase membrane showed
  complete rejection of dyes bigger than 1.2 nm in size
  \cite{zhou_supported_2005}.

  The H\textsubscript{II} phase pore geometry is better for transport than the
  Q\textsubscript{I} phase. The H\textsubscript{II} phase forms at room
  temperature in the presence of c.a.~10 wt\% water and consists of hexagonally
  packed, hydrophilic pore columns\cite{smith_ordered_1997}. In the absence of
  water, neat monomer will form the same hexagonal columnar structure which, in
  literature, has been referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014} (See Figure~\ref{fig:hII_phase}).  The most
  promising Q\textsubscript{I} phase for membrane applications forms at 70\degree
  C when monomer is mixed with c.a.~20 wt\%
  glycerol\cite{carter_glycerol-based_2012}. Q\textsubscript{I} phase membranes
  consist of a tortuous network of three dimensionally interconnected pores that
  prevent optimal through-plane transport. The densely packed, non-tortuous and
  uniform sized pores of H\textsubscript{II} phase membranes represents the ideal
  geometry for achieving high solute flux\cite{matyka_tortuosity-porosity_2008}.
  Despite the promise of the H\textsubscript{II} phase, the hexagonally packed
  liquid crystalline domains, formed during self-assembly of NA-GA3C11 monomers, are
  isotropically aligned which is detrimental to membrane performance.

@@

%MRS5: not a strong thesis.  Why has it been revived?  The reasons for that would be a better thesis.
  Recently, researchers have learned how to macroscopically align the hexagonal
  domains which has revived research into H\textsubscript{II} phase LLC membranes.
%  Research into the H\textsubscript{II} phase of polymerized LLC membranes has
%  been revived in recent years. 
%  During early stages of exploration, hexagonal
%  mesophases formed by Na-GA3C11 (Figure~\ref{fig:assembly}) could not be
%  macroscopically aligned, resulting in low flux membranes, and no clear route
%  towards scalable and economical filtration. 
%  For this reason, research efforts
  Previously, research efforts were focused on the Q\textsubscript{I} phase,
  whose geometry does not require alignment. In 2014, Feng et al.~showed that one
  can align Col\textsubscript{h} hexagonal domains using a magnetic field with
  subsequent cross-linking to lock the structure in place
  \cite{feng_scalable_2014}. In 2016, Feng et al. showed that one could obtain
  the same result using a second technique termed soft confinement
  \cite{feng_thin_2016}. Current efforts are focused on extending the method to
  the H\textsubscript{II} phase and characterizing the performance of these newly
  aligned systems.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to Na-GA3C11 with some characterization done after minor structural
  modifications. Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing
  \cite{resel_structural_2000}. In a later study of rejection performance, it was
  shown that membranes formed by cross-linked Na-GA3C11 in the
  H\textsubscript{II} phase cannot separate solutes less than 1.2 nm in diameter
  because the pores are too large \cite{zhou_supported_2005}. We do not yet
  understand how to controllably reduce the effective pore size or how to tune
  the chemical environment in the nanopores of this or related materials for
  small molecule separations. The only source of predictive modeling for
  LLC systems have been macroscopic models that likely do not adequately describe
  transport at these length scales \cite{hatakeyama_water_2011}. It will be
  challenging to efficiently narrow down the large design space in a laboratory
  setting without a robust model.

@@

  Here we develop a model of the Col\textsubscript{h} assembly formed by
  Na-GA3C11. Compared to the H\textsubscript{II} phase, the Col\textsubscript{h}
  phase is a simpler starting point. The system has no water which will allow us
  to simulate longer timescales, and there exists detailed experimental
  characterization of the fully aligned state, including 2D wide-angle X-ray
  scattering (WAXS) patterns which are useful for reconstructing structural data.
  We anticipate using this Col\textsubscript{h} model in order to build an 
  H\textsubscript{II} model in the future.
  %The two phases appear to share similar structural characteristics since the
  %pore spacings in each system are in close agreement
  %\cite{feng_thin_2016,resel_structural_2000}.

  In order to appropriately model transport in these ordered, nanoporous
  organic systems, we must first gain a thorough understanding of the nanoscopic
  structure of LLC polymer membranes. Our current level of understanding about
  the structure at the nanoscale is limited. Using X-ray diffraction and TEM data
  we can infer that monomers aggregate into hexagonally packed columns with their
  head groups facing towards the column center. There remain several key questions
  that we will investigate in order to expand this understanding:
